Lesson 4
========================================================
```{r Things to run at the start of every session}
library(dplyr)
library(ggplot2)

```

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
#qplot(x=age, y=friend_count, data = pf)
#qplot(age,friend_count, data=pf)

ggplot(aes(x=age, y=friend_count), data=pf) + geom_point()
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:
Add one layer at a time when building plots for debugging purposes

```{r ggplot Syntax}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point() + 
  xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes:
Where points are on top of one another, making it hard to determine many things, such as frequency

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha = 1/20) + # alpha makes it so it takes 20 points to be one black dot
  xlim(13,90) 

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20) + # alpha makes it so it takes 20 points to be one black dot
  xlim(13,90) +
  coord_trans(y='sqrt')
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
# for geom_jitter
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y='sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(age, friendships_initiated), data=pf) +
  geom_point(alpha = 1/25,position = position_jitter(h=0)) +
  xlim(13,90) +
  coord_trans(y="sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:
Think about the limitations of each data point and how it can be bound in the axes to show people
***

### Conditional Means
Notes:
Sometimes you want to see how the mean or median of a variable varies with another variable.
Ex. How does the average friend count vary over age?

```{r Conditional Means}
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n= n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)
```

```{r Alternative way to write above code}
pf.fc_by_age <- pf %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = meadian(friend_count),
            n= n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
```


Create your plot!

```{r Conditional Means Plot}
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:


```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0),
             color = "orange") +
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat="summary", fun.y=quantile, fun.args = (probs = 0.1),
            linetype= 2, color="blue") +
  geom_line(stat="summary", fun.y = median, color = "blue") + 
  geom_line(stat = "summary", fun.y = quantile, fun.args = (probs = 0.9), 
            linetype= 6, color="blue") +
  coord_cartesian(xlim=c(13,30), ylim=c(0,1000))
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

