Lesson 3
========================================================

***

### What to Do First?
Notes:
Understand variables central to your analysis
***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv("pseudo_facebook.tsv", sep = '\t')
# read in the data and indicate that the separator is a tab, not a comma
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:


```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
qplot(x=dob_day, data=pf) + 
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:
Breaks histogram with another data point.
In this case, break all birthdays down into months. 

facet_wrap(forumla)
facet_wrap(~variable)
facet_grid(verticle~horizontal) - generally more useful for 2 or more variables

```{r Faceting}
qplot(x=dob_day, data=pf) + 
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:
Outliers can have many causes.
Outliers can be accurate data!
Sometimes, they can be bad data or limitations of your data. 

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count, data=pf)
```

#### How is this plot similar to Moira's first plot?
Response:
This data is long tailed data. The bulk of our data is <1000 friend counts
***

### Limiting the Axes
Notes:
xlim = c(0,1000) limits x axis from 0 to 1000 to let you get a better view of the data points you want to see.

```{r Limiting the Axes}
#qplot(x=friend_count, data=pf, xlim=c(0,1000))

qplot(x=friend_count, data=subset(pf, !is.na(gender)), binwidth=25) + 
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50))+
  facet_wrap(~gender)
```

### Exploring with Bin Width
Notes:
breaks = seq(starting point, ending end, interval space)

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:
NA values are missing values
!is.na()
We want to use !is.na(gender) because it'll omit any NAs in the dataset. 

```{r Omitting NA Values}
qplot(x=friend_count, data=subset(pf, !is.na(gender)), binwidth=10) + 
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50))+
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:
table() shows counts of variables. In our case, total males and total females

by(variable, categorical variable, function)
by(pf$friend_count, pf$gender, summary)

```{r Statistics \'by\' Gender}
table(pf$gender) # to see if there are more men than women
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:


```{r Tenure}
qplot(x=tenure/356, data=pf, binwidth=0.25,
      color=I('black'), fill=I("#099DD9")) + # color typically comes in second line
  scale_x_continuous(breaks=1,7,1)
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x=tenure/356, data=pf,
      color=I('black'), fill=I("#F79420")) + # color typically comes in second line
  scale_x_continuous(breaks=seq(0,7,1), lim= c(0,7))
```

***

### Labeling Plots
Notes:
label with 
xlab
ylab

These go in the first function, not any + afterwards

You should focus on SCALING and LIMITS on axes
```{r Labeling Plots}
qplot(x=tenure/356, data=pf,
      xlab = "Number of years using Facebook",
      ylab = "Number of users in sample",
      color=I('black'), fill=I("#F79420")) + # color typically comes in second line
  scale_x_continuous(breaks=seq(0,7,1), lim= c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data=pf, binwidth=1,
      color=I('black'), fill=I("#F79420")) + 
  scale_x_continuous(breaks=seq(13,113,5))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
transform data to see patterns more clearly.
This means like log10, log2, or square root.
This is to turn it into a more normal distribution.
```{r}
qplot(x=friend_count, data=pf)

summary(log10(pf$friend_count+1)) # done in orders of 10 
```

```{r Friend Count Transformations}
p1 <- qplot(x=friend_count, data=pf)
p2 <- qplot(x=log10(friend_count+1), data=pf)
p3 <- qplot(x=sqrt(friend_count), data=pf)
grid.arrange(p1,p2,p3, ncol=2)
```

The log10 transformation is much better than the other two. The idea is to try to get this into a normal distribution.

```{r Create the plot with ggplot}
p1 <- ggplot(aes(x=friend_count), data=pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1,p2,p3, ncol=2)

```

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), 
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 600),
                     breaks = seq(0, 1000, 100))
```

```{r Frquency Polygons for www_likes}
#ggplot(aes(x=www_likes), data = pf, geom_freqpoly())

qplot(x=www_likes, data=subset(pf, !is.na(gender)), 
      geom="freqpoly", color=gender)+
  scale_x_continuous()+
  scale_x_log10()

```


***

### Likes on the Web
Notes:

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
```


***

### Box Plots
Notes:
Categorizing variable is X-Axis
Continuous variable is Y-Axis
ylim and scale_y_continuous removes data points which changes the box plot. 

So for box plots, use coord_cortesian
```{r Box Plots}
qplot(x= gender, y=friend_count,
      data = subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim=c(0,1000))
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x= gender, y=friend_count,
      data = subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim=c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x=gender, y=friendships_initiated,
      data = subset(pf, !is.na(gender)),
      geom= "boxplot") + 
  coord_cartesian(ylim=c(0, 200))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:
Some ways to transform a data point is binary such as whether someone uses a feature or not.
```{r Getting Logical}
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```


```{r Getting Logical Quiz}
(sum(pf$mobile_check_in == 1)) / length(pf$mobile_check_in) * 100

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!