Lesson 5
========================================================
```{r Things to run first}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(reshape2)
pf = read.csv("pseudo_facebook.tsv", sep = "\t")
yo <- read.csv("yogurt.csv")
library(GGally)
```


### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram() 
```

```{r}
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(as.numeric(friend_count)),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)
```


***

### Plotting Conditional Summaries
Notes:

```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = "summary", fun.y = median)
```


```{r Plotting Conditional Summaries}
ggplot(aes(age, median_friend_count), data = pf.fc_by_age_gender) + 
  geom_line(aes( color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
```

```{r}
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = "median_friend_count")

head(pf.fc_by_age_gender.wide)
```
```{r}
pf.fc_by_age_gender.test <- melt(pf.fc_by_age_gender.wide, id.vars = c("age, gender"))

head(pf.fc_by_age_gender.test)
```



***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(age, female/male), data = pf.fc_by_age_gender.wide) +
  geom_line() + 
  geom_hline(yintercept = 1, linetype = 2, alpha = 0.3)
```

***

### Third Quantitative Variable
Notes:


```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, 
                          c(2004, 2009, 2011, 2012, 2014))

table(year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA= 'ifany')
```

```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = "summary", fun.y = median)
```

```{r}
ggplot(aes(age, friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket),
            stat = "summary", fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(age, friend_count), 
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket),
            stat = "summary", fun.y = mean) + 
  geom_line(stat = "summary", fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
friending_rate <- subset(pf, tenure > 0)

summary(friending_rate$friend_count/friending_rate$tenure)
```
```{r}
with(subset(pf, tenure > 0), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:
tenure and friend count as a rate
how many friends does a user have for each day they started using service

What is the median friend rate?
0.2205
What is the maximum friend rate?
417
```{r Friendships Initiated}
ggplot(aes(tenure, friendships_initiated/tenure),
       data = subset(pf, tenure > 0)) + 
  geom_line(aes(color = year_joined.bucket),
            stat = "summary", fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```
```{r}
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
```
```{r}
summary(yo)
names(yo)
yo$id <- factor(yo$id)
```
```{r}
ggplot(aes(price), data = yo) + 
  geom_histogram()
```


***

### Number of Purchases
Notes:

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada +
                             plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(time, price), data = yo) + 
  geom_jitter(alpha = 1/5, shape = 20)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# set the seed for reproducible results

set.seed(4320)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(time, price),
       data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~ id) +
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)
```
```{r}
set.seed(6000)
ggplot(aes(time, all.purchases),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(aes(size = price), pch = 10)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
install.packages("GGally")
```
```{r}
library(GGally)
```

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

